Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Remove EqDocs submodule
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@36 db7c848b-7755-4df2-b763-8796797afa52
add
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@37 db7c848b-7755-4df2-b763-8796797afa52
Add
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@38 db7c848b-7755-4df2-b763-8796797afa52

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Stefan Eilemann\thanks{email: eilemann@gmail.com} \\ %
\and Renato Pajarola\thanks{email: pajarola@acm.org}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
  \begin{abstract}
    Developing complex, real world applications which utilize multiple GPUs and
    computers for 3D interactive rendering tasks can be a daunting task, in
    particular since it requires expertise in distributed systems and parallel
    rendering in addition to the application domain. We present a mature
    parallel rendering framework which provides a large set of features,
    algorithms and system integration for a wide range of real-world research
    and industry applications. Based on the basic architecture and
    implementation of the Equalizer parallel rendering framework presented in
    \cite{EMP:09}, we show how the generic algorithms are integrated in the
    framework to help application development in many different domains.
  \end{abstract}

@@

Highlight integration of many features into a general framework vs one-off
research implemenation of one algorithm.

@@

In~\cite{EMP:09}, we presented the Equalizer parallel rendering framework. This
paper summarized the work in parallel rendering up to 2009. An extensive
Programming and User Guide provides in-depth documentation on using and
programming Equalizer~\cite{Eilemann:13}. In the following we assume these two
publications and their references as the basis, and will present the related
work published since then.

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention since 2009. While some
commercial implementations such as TechViz and MechDyne Conduit continue to
exist, on the research side only ClusterGL~\cite{NHM:11} has been
presented. ClusterGL employs the same approach as Chromium, but delivers a
significantly faster implementation of transparent OpenGL interception and
distribution for parallel rendering. CGLX~\cite{DK:11} tries to bring parallel
execution transparently to OpenGL applications, by emulating the GLUT API and
intercepting certain OpenGL calls. In contrast to frameworks like Chromium, CGLX
follows the distribution application approach, whichworks transparently for
trivial applications, but quickly requires the application developer to address
the complexities of a distributed application when mutable application state
needs to be synchronized across processes.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention. Sage~\cite{Sage} and Sage~2~\cite{Sage2}
have received significant attention. Sage is build entirely around the concept
of a shared framebuffer where all content windows are separate applications
using pixel streaming. It is no longer actively supported. Sage 2 is a complete,
browser-centric reimplementation where each application is a web application
distributed across browser instances. DisplayCluster~\cite{DC}, and its
continuation Tide~\cite{tide}, also implement the shared framebuffer concept of
Sage, but provide a few native content applications integrated into the display
servers.

Equalizer itself has received significant attention within the research
community. Various algorithms to improve the parallel rendering performance have
been proposed: compression and region of interest during
compositing~\cite{MEP:10}; load-balancing resources for multi-display
installations~\cite{EEP:11}; asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12}; as well as work queueing~\cite{SPEP:16}.

Various applications and frameworks have used Equalizer for new research in
visualization, such as the RTNeuron~\cite{HBBES:13}, an OpenGL renderer for
visualizing simulations of neucortical brain regions; Omegalib~\cite{Omegalib},
extending Equalizer for the hybrid 2D/3D environments such as the CAVE2;
visualization for remote sensing data~\cite{LK:09}, bino.

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{EqDocs/Developer/ProgrammingGuide/images/netObject}
  \caption{\label{fNetObject}Communication between two Collage objects}
\end{figure}
\fig{fNetObject} illustrates this architecture.

>>>>

%% Author and Affiliation (multiple authors with multiple affiliations)
\author{Stefan Eilemann\thanks{email: eilemann@gmail.com} \\ %
\and Renato Pajarola~\IEEEmembership{Senior Member,~IEEE}\thanks{email: pajarola@acm.org}}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEtitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\IEEEtitleabstractindextext{%
  \begin{abstract}
    Developing complex, real world graphics applications which utilize multiple GPUs and
    computers for interactive 3D rendering tasks can be a daunting task, in
    particular since it requires expertise in distributed systems and parallel
    rendering in addition to the application domain itself. We present a mature
    parallel rendering framework which provides a large set of features,
    algorithms and system integration for a wide range of real-world research
    and industry applications. Based on the basic architecture and
    implementation of the Equalizer parallel rendering framework presented in
    \cite{EMP:09}, we show how a variety of generic algorithms are integrated in the
    framework to help application development in many different domains, and we show
    a number of applications highlighting the diverse aspects and use cases of Equalizer.
  \end{abstract}

@@

Highlight integration of many features into a general framework vs one-off
research implementation of one algorithm. Also point out the various and very
diverse examples in Section 7, highlighting the variety of application and implementation
scenarios that can benefit from Equalizer.

@@

In~\cite{EMP:09}, we presented the initial {\em Equalizer} parallel rendering framework,
which summarized the work in parallel rendering up to 2009. An extensive
Programming and User Guide provides in-depth documentation on using and
programming Equalizer~\cite{Eilemann:13}. In the following we assume these two
publications and their references as a starting point, and focus on the related
work published since then.

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention since 2009. While some
commercial implementations such as TechViz and MechDyne Conduit continue to
exist, on the research side only ClusterGL~\cite{NHM:11} has been
presented. ClusterGL employs the same approach as Chromium, but delivers a
significantly faster implementation of transparent OpenGL interception and
distribution for parallel rendering. CGLX~\cite{DK:11} tries to bring parallel
execution transparently to OpenGL applications, by emulating the GLUT API and
intercepting certain OpenGL calls. In contrast to frameworks like Chromium, CGLX
follows the distributed application approach, which works transparently for
trivial applications, but quickly requires the application developer to address
the complexities of a distributed application when mutable application states
need to be synchronized across processes.\FIXME{do we need to state how Equalizer
is different from the new ClusterGL and CGLX?}

On the other hand, software for driving and interacting with tiled display walls
has received significant attention, including Sage~\cite{Sage} and Sage~2~\cite{Sage2}
in particular. Sage was built entirely around the concept
of a shared framebuffer where all content windows are separate applications
using pixel streaming but is no longer actively supported. Sage 2 is a complete,
browser-centric reimplementation where each application is a web application
distributed across browser instances. DisplayCluster~\cite{DC}, and its
continuation Tide~\cite{tide}, also implement the shared framebuffer concept of
Sage, but provide a few native content applications integrated into the display
servers.\FIXME{also here, maybe a differentiation to Sage2 and DC is needed,
showing the advantages of EQ}

Equalizer itself has received significant attention within the research
community. Various algorithms to improve the parallel rendering performance have
been proposed: compression and region of interest during
compositing~\cite{MEP:10}; load-balancing resources for multi-display
installations~\cite{EEP:11}; asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12}; as well as work queueing~\cite{SPEP:16}.
Additionally, complex large scale and out-of-core multiresolution
rendering approaches have been parallelized and implemented with
Equalizer~\cite{GMBP:10, GEMPG:13}, demonstrating the feasibility of
the framework to be used with complex rendering algorithms and 3D model
representations.

Furthermore, various applications and frameworks have used Equalizer for new research in
visualization, such as the RTNeuron~\cite{HBBES:13}, an OpenGL renderer for
visualizing simulations of neurocortical brain regions; Omegalib~\cite{Omegalib},
extending Equalizer for the hybrid 2D/3D environments such as the CAVE2;
visualization for remote sensing data~\cite{LK:09}, bino.

@@

\begin{figure}[ht]\center
  \includegraphics[width=\columnwidth]{images/netObject}
  \caption{\label{fNetObject}Communication between two Collage objects}
\end{figure}
\fig{fNetObject} illustrates this architecture.
